---
title: "ABA analysis"
author: "Alonso Serrano"
date: "7/10/2019"
updated: "21/01/2020"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir="/mnt/picea/home/aserrano/drought")
```

# Introduction
Plot ABA related genes, their infomap clusters (if any) and general information

## Prerequisites

Dataset with variance stabilization transformation applied
Dataset metadata
Infomap clusters
Table with DE gene results


## Setup
Load the modules and auxiliary scripts 
```{r ini}
library(here)
library(dplyr)
#library(plyr)
source(here("Rtoolbox/src/infomapTools.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))
source(here("Rtoolbox/src/plotEigenGene.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))

projectDir <- "/mnt/picea/projects/spruce/vhurry/drought-stress-combined"
deDir <- file.path(projectDir, "de")
networkDir <- file.path(projectDir, "network")

```

Load the drought vst data obtained from Julia's analysis
```{r loadData, include=F}
load(file.path(deDir, "combinedData.RData"))
```

Load the drought vst data obtained from Julia's analysis
```{r, eval=FALSE}
combinedData <- read.table(file.path(deDir,"vsdDrought.tsv"), 
                           header=T, row.names = 1, sep='\t')
combinedData <- as.matrix(combinedData)
```
```{r metadata}
meta <- read.csv(here("doc","droughtcombined.csv"))
meta <- meta[-15,]
meta$Level <- factor(meta$Level, levels = c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate") )

meta$days <- plyr::mapvalues(meta$Level, 
          from=c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate"), 
          to=c(0,2,4,5,13,18,21,25))

meta$days <- as.numeric(as.character(meta$days))

meta$points <- plyr::mapvalues(meta$Level, 
          from=c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate"), 
          to=c("Control", "2 Days", "4 Days", "5 Days",
                                            "13 Days", "18 Days", "21 Days","Re-irrigation"))

```

```{r loadBigTable, include=FALSE}
bigTable <- read.table(file.path(projectDir, "results", "table1_infomap_stats.tsv"),
                       sep='\t', header=T, stringsAsFactors = F, comment.char = "",
                       quote="")
bigTable[is.na(bigTable$infomapCluster),]$infomapCluster <- "noCluster"
```

```{r loadABAlist}
# 
# table1 <- read.table("~/Git/drought-combined-network/doc/table1.tsv", 
#                      sep='\t', comment.char = "", header=T,
#                      stringsAsFactors=FALSE, quote="")

ABAlistTable <- read.table(here("doc","ABAsublist.tsv"), sep='\t', 
                      header = T,fill=T, quote="", stringsAsFactors = F)

genesOI <- ABAlistTable$Congenie.Norway.spruce.ID

ABAlistTemp <- bigTable[bigTable$Congenie.Norway.spruce.ID %in% genesOI,]

ABAlist <- data.frame(Congenie.Norway.spruce.ID = ABAlistTemp$Congenie.Norway.spruce.ID, 
                      Tair.Arabidopsis.ID = ABAlistTemp$Tair.Arabidopsis.ID,
                      Tairinfo = ABAlistTemp$Tairinfo,
                      Nickname = ABAlistTemp$Nickname,
                      infomapCluster = ABAlistTemp$infomapCluster,
                      stringsAsFactors = F) 
ABAlist <- distinct(ABAlist)
```


```{r loadInfomap}
infomapClusters <- read.table(file.path(networkDir, "cluster/InfomapClusters.tsv"),
                              header=T,stringsAsFactors = F)
names(infomapClusters) <- c("gene", "infomapCluster")
```

```{r mergeTables, eval=FALSE}
# ABAlist <- left_join(ABAlist , infomapClusters, by = "Congenie.Norway.spruce.ID")
# ABAlist[is.na(ABAlist$infomapCluster),]$infomapCluster <- "noCluster"
```

# Cluster index
```{r addIndex, message=FALSE, echo=F, results="asis"}

lapply(unique(ABAlist$infomapCluster), function (clus){
  cat(paste0("<a href=\"#",clus,"\">",clus,"</a>"))
  cat("\n\n")
})

```


```{r, message=FALSE, results="asis"}

narrowGeneList <- read.table(here("src","narrowGeneList.tsv"), quote="\"", comment.char="", stringsAsFactors=FALSE)[,1]

miniPrint <- function(genes) {
  for(gen in genes) {
    #if (gen %in% narrowGeneList) {
      
      thisData <- ABAlist[ABAlist$Congenie.Norway.spruce.ID==gen,]
      cat("<H3>Gene</H3>")
      cat("\n\n")
      cat(gen)
      cat("\n\n")
      for(id in thisData$Tair.Arabidopsis.ID){
        thisRow <- thisData[thisData$Tair.Arabidopsis.ID==id,]
        cat("<H5>Tair Arabidopsis ID</H5>")
        cat("\n\n")
        cat(thisRow$Tair.Arabidopsis.ID)
        cat("<H5>Tairinfo</H5>")
        cat("\n\n")
        cat(thisRow$Tairinfo)
        cat("\n\n")
        cat("<H5>Nickname</H5>")
        cat("\n\n")
        cat(thisRow$Nickname)
        cat("\n\n")
      }
      cat("<H5>Plot</H5>")
      cat("\n\n")
      if(gen %in% colnames(combinedData)){
        print(plotEigengene(combinedData, gen, meta$Type, meta$days, timeUnits = "Days",    
                noLegend = F, legendTitle="Tissue"))
        cat("\n\n")
      } else {
        cat("NO DATA AVAILABLE")
        cat("\n\n")
      }
      cat("<hr>")
      cat("<hr>")
      #}
    }
    
}

#a full list
for (cluster in unique(ABAlist$infomapCluster)) {
#a narrow list
#for (cluster in c("noCluster", "Cluster1:2", "Cluster1:10", "Cluster1:5", "Cluster1:7")) {  
  cat(paste0("<a name=\"",cluster,"\"></a><H2>",cluster,"</H2>"))
  cat("\n\n")
  
  if(cluster!="noCluster") {
    fullCluster <- infomapClusters[infomapClusters$infomapCluster==cluster,]$gene
    print(plotEigengene(combinedData, fullCluster, meta$Type, meta$days, timeUnits = "Days",    
                noLegend = F, legendTitle="Tissue"))
    cat("\n\n")
  }
  
  temp <- ABAlist[ABAlist$infomapCluster==cluster,]$Congenie.Norway.spruce.ID
  
  if(cluster=="noCluster") {
    temp <- intersect(temp, narrowGeneList)
  }
  miniPrint(temp)
  cat("\n\n")
  cat("<hr>")

}
```


