---
title: "drought Network Analysis"
author: "Alonso Serrano"
date: "8/14/2019"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir="/mnt/picea/home/aserrano/drought")
```

# Introduction
This file contains the script and results for the clustering and enrichment of the drought network
from Julia Hass drought analysis. 

## Prerequisites

Dataset with variance stabilization transformation applied
Dataset metadata
Aggregated network file from Seidr
Backbone file obtained from the aggregated network file


## Setup
First we load the modules and auxiliary scripts 
```{r ini}
library(here)
library(RLinuxModules)
library(data.table)
module("load bioinfo-tools seidr-devel")
module("load bioinfo-tools InfoMap")

source(here("Rtoolbox/src/infomapTools.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))
source(here("Rtoolbox/src/plotEigenGene.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))

projectDir <- "/mnt/picea/projects/spruce/vhurry/drought-stress-combined/"
deDir <- paste0(projectDir, "de/")
networkDir <- paste0(projectDir, "network/")

```

```{r loadData, include=F}
load(paste0(deDir, "combinedData.RData"))
```

Load the drought vst data obtained from Julia's analysis
```{r, eval=FALSE}
combinedData <- read.table(paste0(deDir,"vsdDrought.tsv"), header=T, row.names = 1, sep='\t')
combinedData <- as.matrix(combinedData)
```

Load the metadata and delete sample 15, to be coherent with the paper.
On the droughtAUC.R file, we have seen that the best backbone we can use is the 1%. 
That is on the backbone-1-percent.sf file and the algorithm that we will use is the irp_score.
Now we define those parameters, as well as the edge file and the names of the files that will come after Infomap.

```{r load}
meta <- read.csv(here("doc","droughtcombined.csv"))
meta <- meta[-15,]
#seidrExe <- ("/mnt/picea/home/bastian/Git/seidr-devel/build/seidr ")
seidrExe <- ("seidr ")
backboneFile <- paste0(networkDir, "results/backbone/backbone-1-percent.sf")
clusterFolder <- paste0(networkDir, "cluster/")
dir.create(clusterFolder, showWarnings = FALSE)
algo <- "irp_score"
edgeFile <- paste0(clusterFolder,"edgeList.txt")
edgeIndexFile <- paste0(clusterFolder,"edgeIndexList.txt")
treeFile <- paste0(clusterFolder,"edgeIndexList.tree")

```

# Infomap clustering
First we use Seidr reheader function to drop not connected nodes and improve efficiency
```{r reheader, results=F, message='HIDE'}
system(paste(seidrExe, "reheader" , backboneFile), intern=TRUE)

```

We execute seidr view to extract the information of source node, target node and the score of irp
```{r include=FALSE}

headResult <- system(paste(seidrExe, "view", backboneFile, "-c -d $'\t' ", "| head -n 1"), intern=TRUE)
headResult <- unlist(strsplit(headResult, "\t"))

algoIndex <- grep("irp_score", headResult)
system(paste0(seidrExe, "view ", backboneFile, "  -d $'\t' | cut -f 1,2,",algoIndex, " > ",edgeFile), intern=TRUE)
system(paste0(seidrExe, "view ", backboneFile, " -N -d $'\t' | cut -f 1,2,",algoIndex," >",edgeIndexFile), intern=TRUE)

```

We will execute Infomap with a markov-time value of 1.4, because that value matches our goal to 
keeps between 50-60% of the genes in the top 20 clusters.
```{r infomap, results=F, message='HIDE'}
markovTime <- 1.4
system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0(seidrExe, " resolve -s ", backboneFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  
```

We can get some information about the quality of the clusters
```{r qa1}
print(paste("% of genes in the top 20 clusters in Level 1:", clusterQA(infomapTable)))
print(paste("% of genes in the top 20 clusters in Level 2:", clusterQA(infomapTable, level='Level2')))

```

We select those clusters that at least has 40 genes. Our goal is to obtain between 25-35 clusters
```{r}
selectedClusters <- getClusterByMinSize(infomapTable, level = 'Level2',min=40)

topClusters <- getClusters(infomapTable, "Level2", numberOfClusters=selectedClusters)
```

Then we save the results of clustering for network visualization
```{r save4Cytoscape, eval=F}
save4Cytoscape(topClusters, file=paste0(clusterFolder,"InfomapClusters.tsv"))
save4Cytoscape(topClusters, file=here("results", "InfomapClusters.tsv"))
```

# Enrichment

We use as background all genes that at least have 1 count in any condition of the dataset, we only remove 
genes with absolute zero expression. this information was extracted after the DE analysis but before 
removing the zero MAD genes
```{r gofer}
background <- read.table(paste0(deDir, "background.txt"), stringsAsFactors = F, header=F)[[1]]
```

We use gofer2 to obtain all the enrichments for each cluster. GO, pfam, KEGG and Mapman.
By default we filter all the terms above 0.05 of pAdj value
```{r, eval=FALSE}
clusterEnr <- enrichClusters(topClusters, background = background)
save(paste0(clusterFolder, "clusterEnr.RData"))
```

```{r loadEnr, include=FALSE}
load(paste0(clusterFolder, "clusterEnr.RData"))
```

We save the enrichment data to XLS files for future uses.
```{r save, eval=FALSE}
enr2xls(clusterEnr, filePrefix = paste0(clusterFolder,"drought" ) )
```

# Plots

In order to get a proper plot, we need to specify the correct order of levels
```{r relevel}
meta$Level <- factor(meta$Level, levels = c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate") )
```

We plot the top 15 clusters expression profiles and GO biological process enrichment treemap.
```{r plot, message=FALSE, results="asis"}
for (clus in 1:15) {
  
  clusName <- names(topClusters)[clus]
  cat(paste0("<H3>", clusName, "","</H3>"))
  cat("\n\n")
  print(plotEigengene(combinedData, topClusters[[clusName]],                      
                meta$Type, meta$Level, timeUnits = "Level",    
                noLegend = F, legendTitle="Tissue",
                title = paste(clusName,"Genes:",length(topClusters[[clusName]]) )))
  cat("\n")
  enr <- clusterEnr[[clusName]][['go']]
  if (!is.null(enr) && any(enr$namespace =='BP')) {
    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'go', namespace = 'BP',
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus],
                        title = paste(clusName," GO enrichment for BP"))
  }
  cat("\n")
  cat("<hr>")
  enr <- clusterEnr[[clusName]][['mapman']]

  if (!is.null(enr)) {
    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'mapman',
                      clusterColor=clusterTreemapColors[clus],
                      clusterText=clusterTreemapText[clus],
                      title = paste(clusName," MAPMAN enrichment"))
  }
  cat("\n")
  cat("<hr>")
  
}
```


# Export results
We can save each plot as a PNG for 
```{r, eval=FALSE}
for (clus in 1:length(topClusters)) {
  print(names(topClusters)[clus])
  thisCluster <- names(topClusters)[clus]
  png(paste0("expressionProfiles/",thisCluster,".png" ), width = 1482, height=571 )
  print(plotEigengene(combinedData, topClusters[[thisCluster]], 
                      meta$Type, 
                      meta$Level, 
                      timeUnits = "Level",    
                      noLegend = F,
                      legendTitle="Tissue",
                      title = paste(thisCluster,"Genes:",length(topClusters[[thisCluster]]) )))
  dev.off()
  
}

```

```{r saveEnr, eval=FALSE}
for (clus in 1:15) {
 
  clusName <- names(topClusters)[clus]
   print(clusName)
  png(paste0("clusterTreemaps/",clusName,"_enr_","go_bp",".png" ), width = 1482, height=571 )
  
  enr <- clusterEnr[[clusName]][['go']]
  if (!is.null(enr)) {
    

    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'go', 
                        namespace = 'BP',
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus])
  }
  dev.off()
}
```

# Session information
Seidr version 0.11.0

Infomap version 0.19.26


```{r}
sessionInfo()
```