---
title: "Drought stress network"
author: "Alonso Serrano"
date: "8/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This file contains the script and results from the backbone analysis of the drought combined network.

## Prerequisites
Backbone files obtained from the aggregated network file
ROC files obtained from the backbone files

## Libraries
Setting up folders
```{r}
projectDir <- "/mnt/picea/projects/spruce/vhurry/drought-stress-combined/"
rocDir <- paste0(projectDir,"network/results/roc/")
```

Load required libraries
```{r libraries, message=FALSE}
library(ggplot2)
library(gplots)
library(LSD)
library(pander)
library(pracma)
library(RColorBrewer)
library(viridis)

```

Color palette
```{r palette}
hpal <- colorRampPalette(c("blue","white","red"))(100)
cols <- brewer.pal(12,"Paired")
mar <- par("mar")
```
# Data

```{r data}
AUCs <- sapply(dir(rocDir, pattern=".*_roc.tsv$",full.names=TRUE), function(f){
  
  algo <- sub("[^_]+_","",sub("_roc\\.tsv","",f))       
  dat <- read.delim(f, header=FALSE, skip=1, col.names = c("TP","FP","PR"))
  head <- scan(f, nmax = 3,what = "character",sep="\t")
  message(sprintf("Found %s GS edges out of %s edges",head[2],sum(as.integer(head[2:3]))))
  auc <- round(trapz(dat[,2],dat[,1]),digits=3)
  
  plot(dat[,2],dat[,1],type="l",main=sprintf("%s (AUC = %s)",algo,auc),
      xlab="False Positive Rate",ylab="True Positive Rate",
      sub=sprintf("%s Gold Standard edges out of %s edges\n%s",head[2],sum(as.integer(head[2:3])),
                  basename(f)))
  abline(0,1,lty=2)
  message(sprintf("The AUC is %s",auc))
  
  return(c(sub("_.*","",basename(f)),algo,auc))
})

tab <- split.data.frame(t(AUCs[2:3,]),AUCs[1,])

stopifnot(nrow(unique(t(sapply(tab,function(ta){ta[,1]})))) == 1)

res <- sapply(tab,function(ta){as.numeric(ta[,2])})
rownames(res) <- tab[[1]][,1]
res <- res[,c(2:ncol(res),1)]


```

Re-order the data.
```{r names}
rocord <- c("backbone-1-percent", "backbone-2-percent", "backbone-3-percent",
            "backbone-4-percent", "backbone-5-percent", "backbone-6-percent", 
            "backbone-7-percent", "backbone-8-percent", "backbone-9-percent", 
            "backbone-10-percent")
```
Create a heatmap of the AUC values for each network and for each algorithm.
```{r heatmap}
heatmap.2(res,
          trace="none",
          margins = c(10.1,7.1),
          col = hpal)

heatmap.2(as.matrix(res[,rocord]), 
          Colv = FALSE,
          dendrogram = "row",
          trace="none",
          margins = c(10.1,7.1),
          col = hpal)
```

Obtain a line-plot from this heatmap.
```{r}
par(mar=c(10.1,3.1,3.1,0.1))
linesplot(res[,rocord],cols = colorRampPalette(c("blue","red"))(20),addboxes = TRUE,las=2,main="AUC distribution")
par(mar=mar)

par(mar=c(10.1,4.1,3.1,2.1))
plot(0,0,ylim=c(0,1),
     xlim=c(1,ncol(res)),type="n",
     xaxt="n",xlab="",ylab="AUC",main="AUC evolution")
axis(1,1:ncol(res),las=2,labels=colnames(res))
sapply(1:nrow(res),function(i){lines(res[i,],col=cols[i],lwd=2,lty=2)})
legend("bottomright",legend = rownames(res),col=cols[1:nrow(res)],lty=1,lwd=2)
par(mar=mar)

pander(res)
write.csv(res[,rocord],file=paste0(rocDir,"AUCs.csv"),quote=FALSE)
```
# Results
Based on the irp overall score and the individual results, we will use the 1% file, for downstream
analysis.


# Session Info
```{r session info, echo=FALSE}
sessionInfo()
```
