---
title: "Drought network analysis"
author: "Alonso Serrano"
date: "`r format(Sys.time(), '%Y %B %d')`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
* 20190814 Creation
* 20200901 Eigengene plot x-axis correct labeling

This file contains the script and results for the clustering and enrichment of the drought network
from Julia Hass drought analysis. 

## Prerequisites

* Dataset with variance stabilization transformation applied  
* Dataset metadata  
* Aggregated network file from Seidr  
* Backbone file obtained from the aggregated network file  

## Setup
First we load the modules and auxiliary scripts 
```{r ini, message=FALSE, warning=FALSE, include=T, echo=T}
library(here)
library(data.table)

# Auxiliary function to load Linux modules
module <- function(Arguments, moduleCmd="/usr/lib/x86_64-linux-gnu/modulecmd.tcl") { 
  rCmds <- system(paste(moduleCmd, "r", Arguments), intern=T) 
  ( eval( parse(text = rCmds) ) )
}
none <- module("load bioinfo-tools seidr-devel")
none <- module("load bioinfo-tools InfoMap")

source(here("Rtoolbox/src/infomapTools.R"))
source(here("Rtoolbox/src/plotEnrichedTreemap.R"))
source(here("Rtoolbox/src/plotEigenGene.R"))

projectDir <- "/mnt/picea/projects/spruce/vhurry/drought-stress-combined"
deDir <- file.path(projectDir, "de")
networkDir <- file.path(projectDir, "network")

```

```{r loadData, include=F}
# we load the stored combinedData for speed up
load(file.path(deDir, "combinedData.RData"))
```

Load the drought vst data obtained from Julia's analysis
```{r, eval=FALSE}
combinedData <- read.table(file.path(deDir, "vsdDrought.tsv"), header=T, row.names = 1, sep='\t')
combinedData <- as.matrix(combinedData)
```

Load the metadata and delete sample 15 as it was identified as an outlier in the
analysis done on drought_pub_scrip1.R  

On the droughtAUC.R file, we have seen that the best backbone we can use is the 1%. 
That is on the backbone-1-percent.sf file and the algorithm that we will use is the irp_score.  

Now we define those parameters, as well as the edge file and the names of the files that will come after Infomap.

```{r load}
meta <- read.csv(here("doc", "droughtcombined.csv"))
meta <- meta[-15,]
backboneFile <- file.path(networkDir, "results/backbone/backbone-1-percent.sf")
clusterFolder <- file.path(networkDir, "cluster")
dir.create(clusterFolder, showWarnings = FALSE)

algo <- "irp_score" # Index score to use
edgeFile <- file.path(clusterFolder, "edgeList.txt")
edgeIndexFile <- file.path(clusterFolder, "edgeIndexList.txt")
treeFile <- file.path(clusterFolder, "edgeIndexList.tree")
resolveFile <- file.path(clusterFolder, "seidrResolve.txt")
```

# Infomap clustering
First we use Seidr reheader function to drop not connected nodes and improve efficiency
```{r reheader, message=FALSE, results=F, eval=F}
system(paste("seidr reheader" , backboneFile), intern=TRUE)
```

We execute seidr view to extract the information of source node, target node and the score of irp
```{r eval=FALSE}
headResult <- system(paste("seidr view", backboneFile, "-c -d $'\t' ", "| head -n 1"), intern=TRUE)
headResult <- unlist(strsplit(headResult, "\t"))

algoIndex <- grep("irp_score", headResult)
system(paste0("seidr view ", backboneFile, " -d $'\t' | cut -f 1,2,",algoIndex, " > ",edgeFile), intern=TRUE)
system(paste0("seidr view ", backboneFile, " -N -d $'\t' | cut -f 1,2,",algoIndex," >",edgeIndexFile), intern=TRUE)
```

We will execute Infomap with a markov-time value of 1.4, because that value matches our goal to 
keep between 50-60% of the genes in the top 20 clusters.
```{r infomap, eval=FALSE, message=FALSE, include=T, results=F}
markovTime <- 1.4
system(paste("Infomap ", edgeIndexFile," -z --markov-time ", markovTime," ", clusterFolder))
infomapRes <- system(paste0("seidr resolve -s ", backboneFile, " ", treeFile), intern=TRUE)
infomapTable <-data.frame(do.call(rbind, strsplit(infomapRes, "\t")))
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2)  
```

```{r include=FALSE}
infomapTable <- read.table(resolveFile, header = T, sep='\t')
infomapTable <- prepareData(infomapTable)
infomapTable$Level1 <- infomapTable$P1
infomapTable$Level2 <- paste0(infomapTable$Level1,":",infomapTable$P2)
infomapTable$Level2 <- ifelse(infomapTable$Level2 %like% "NA", NA, infomapTable$Level2) 
```

We can get some information about the quality of the clusters
```{r qa1}
print(paste("% of genes in the top 20 clusters in Level 1:", clusterQA(infomapTable)))
print(paste("% of genes in the top 20 clusters in Level 2:", clusterQA(infomapTable, level='Level2')))
```

We select those clusters that at least has 40 genes. Our goal is to obtain between 25-35 clusters
```{r}
selectedClusters <- getClusterByMinSize(infomapTable, level = 'Level2',min=40)
topClusters <- getClusters(infomapTable, "Level2", numberOfClusters=selectedClusters)
```

Then we save the results of clustering for network visualization
```{r save4Cytoscape, eval=F}
save4Cytoscape(topClusters, file=file.path(clusterFolder,"InfomapClusters.tsv"))
save4Cytoscape(topClusters, file=here("results", "InfomapClusters.tsv"))
```

```{r include=FALSE}
# since it is already calculated we can safely load the file
read.table
```


# Enrichment

We use as background for the enrichment tests all genes that are differentially expressesed as determined by the analysis in drought_pub_script1.R 
```{r gofer}
background <- read.table(file.path(deDir, "background.txt"), stringsAsFactors = F, header=F)[[1]]
```

We use gofer2 to obtain all the enrichments for each cluster. 
Tested for GO and Mapman enrichments.  
By default, all the results with an adjusted p-value above 0.05 are removed.
```{r, eval=FALSE}
clusterEnr <- enrichClusters(topClusters, background = background)
save(file.path(clusterFolder, "clusterEnr.RData"))
```

```{r loadEnr, include=FALSE}
load(file.path(clusterFolder, "clusterEnr.RData"))
```

We save the enrichment data to XLS files for future uses.
```{r save, eval=FALSE}
enr2xls(clusterEnr, filePrefix = file.path(clusterFolder,"drought") )
```

# Plots

In order to get a proper plot, we need to specify the correct order of levels
```{r relevel}
meta$Level <- factor(meta$Level, levels = c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate") )
meta$days <- plyr::mapvalues(meta$Level, 
          from=c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate"), 
          to=c(0,2,4,5,13,18,21,25))

meta$days <- as.numeric(as.character(meta$days))

meta$points <- plyr::mapvalues(meta$Level, 
          from=c("80%", "60%", "40%", "30%", "30%7d", "Collapse", "C2d", "Rehydrate"), 
          to=c("Control", "2 Days", "4 Days", "5 Days", "13 Days", "18 Days", "21 Days","Re-irrigation"))
```

We plot the top 15 clusters expression profiles and GO biological process enrichment treemap.
```{r plot, message=FALSE, warning=FALSE, results="asis"}
for (clus in 1:15) {
  
  clusName <- names(topClusters)[clus]
  cat(paste0("<H3>", clusName, "","</H3>"))
  cat("\n\n")
  print(plotEigengene(combinedData, topClusters[[clusName]],                      
                meta$Type, meta$days, timeUnits = "Day of Experiment",    
                noLegend = F, legendTitle="Tissue",
                title = paste(clusName,"Genes:",length(topClusters[[clusName]]) )))
  cat("\n")
  enr <- clusterEnr[[clusName]][['go']]
  if (!is.null(enr) && any(enr$namespace =='BP')) {
    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'go', namespace = 'BP',
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus],
                        title = paste(clusName," GO enrichment for BP"))
  }
  cat("\n")
  cat("<hr>")
  enr <- clusterEnr[[clusName]][['mapman']]

  if (!is.null(enr)) {
    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'mapman',
                      clusterColor=clusterTreemapColors[clus],
                      clusterText=clusterTreemapText[clus],
                      title = paste(clusName," MAPMAN enrichment"))
  }
  cat("\n")
  cat("<hr>")
  
}
```


# Export results
We can save each plot as a PNG for 
```{r, eval=FALSE}
for (clus in 1:length(topClusters)) {
  print(names(topClusters)[clus])
  thisCluster <- names(topClusters)[clus]
  png(paste0("expressionProfiles/",thisCluster,".png" ), width = 1482, height=571 )
  print(plotEigengene(combinedData, topClusters[[thisCluster]], 
                      meta$Type, 
                      meta$days, 
                      timeUnits = "Day of Experiment",    
                      noLegend = F,
                      legendTitle="Tissue",
                      title = paste(thisCluster,"Genes:",length(topClusters[[thisCluster]]) )))
  dev.off()
  
}

```

```{r saveEnr, eval=FALSE}
for (clus in 1:15) {
 
  clusName <- names(topClusters)[clus]
   print(clusName)
  png(paste0("clusterTreemaps/",clusName,"_enr_","go_bp",".png" ), width = 1482, height=571 )
  
  enr <- clusterEnr[[clusName]][['go']]
  if (!is.null(enr)) {
    

    plotEnrichedTreemap(clusterEnr[[clusName]], enrichment = 'go', 
                        namespace = 'BP',
                        clusterColor=clusterTreemapColors[clus],
                        clusterText=clusterTreemapText[clus])
  }
  dev.off()
}
```

# Session information
Seidr version 0.11.0

Infomap version 0.19.26


```{r}
sessionInfo()
```